### 2024-12-02 paper/paper.qmd
some graphs and figures

---
title: "Voices in Parliament: Analyzing Interruptions Across Gender in the Australian Political Landscape"
subtitle: "My subtitle if needed"
author: 
  - Hannah Yu
  - Rohan Alexander
thanks: "Code and data are available at: https://github.com/hannahyu07/Hansard_Interruptions"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(here)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(knitr)
library(tibble)
library(gt)
library(patchwork)

```

```{r}
#| include: false
#| warning: false
#| message: false
hansard_corpus <- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))
```

```{r}
#| include: false
#| warning: false
#| message: false
# View the first few rows
head(hansard_corpus)

```

```{r}
#| include: false
#| warning: false
#| message: false
summary(hansard_corpus)
```

```{r}
#| include: false
#| warning: false
#| message: false
# Calculate interruptions and speeches by gender, including rounding the proportion
interruptions_by_gender <- hansard_corpus %>%
  group_by(gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    .groups = 'drop'  # Automatically drop the grouping
  ) %>%
  mutate(
    interruption_proportion = round(total_interruptions / total_speeches, 2)  # Calculate and round the proportion of interruptions
  )

# Calculate total interruptions and speeches for all genders combined, with rounding
interruptions_all <- hansard_corpus %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n()
  ) %>%
  mutate(
    gender = "All",
    interruption_proportion = round(total_interruptions / total_speeches, 2)  # Round the proportion here as well
  )

# Combine the results for specific genders and "All"
final_results <- bind_rows(interruptions_by_gender, interruptions_all)

# Print the results
print(final_results)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Gender-Based Interruption Analysis in Parliamentary Speeches "
#| label: tbl-1

# Create a clean and professional table
final_results %>%
  gt() %>%
  tab_header(
    title = "Proportion of Interruptions by Gender",
  ) %>%
  fmt_percent(
    columns = interruption_proportion,
    decimals = 2
  ) %>%
  cols_label(
    gender = "Gender",
    total_interruptions = "Total Interruptions",
    total_speeches = "Total Speeches",
    interruption_proportion = "Interruption Proportion"
  ) %>%
  tab_style(
    style = cell_borders(sides = "all", color = "gray", weight = px(1)),
    locations = cells_body()
  ) %>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(18),
    heading.subtitle.font.size = px(14),
    data_row.padding = px(5),
    table.border.top.color = "gray",
    table.border.bottom.color = "gray"
  )

```

@tbl-1 provides a clear view of the proportion of interruptions by gender across all speeches in the dataset. The results show that male speakers were interrupted more frequently (21% of speeches) compared to female speakers (18% of speeches). Interestingly, the interruption proportion for unknown gender speakers is significantly lower, at only 11%. Overall, considering all speeches regardless of gender, the interruption proportion stands at 18%. This data suggests that gender may play a role in how frequently speakers are interrupted during parliamentary sessions, with male speakers facing a higher rate of interruption. This could reflect broader dynamics of gender interaction within the parliamentary setting.

```{r}
#| include: false
#| warning: false
#| message: false
#Based on the following data it shows that in fact men seem to be interrupted more on avergae than women in the australian parliment, but this could be changed from the "unknown" category.
# Summarize interruptions by year and gender
interruptions_summary <- hansard_corpus %>%
  group_by(year, gender) %>%
  summarise(
    total_interruptions = sum(as.numeric(interject), na.rm = TRUE),
    .groups = 'drop'
  )

# Summarize total interruptions for all genders combined per year
interruptions_all <- hansard_corpus %>%
  group_by(year) %>%
  summarise(
    total_interruptions = sum(as.numeric(interject), na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    gender = "All"
  )

# Combine gender-specific and "All" categories data
final_data <- bind_rows(interruptions_summary, interruptions_all)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-1
#| fig-cap: "Yearly Interruptions in Australian Parliament by Gender"

ggplot(final_data, aes(x = year, y = total_interruptions, color = gender)) +
  geom_line(size = 1.2) +  # Thicker lines for better visibility
  geom_point(size = 2) +  # Larger points to highlight data
  labs(
    title = "Yearly Interruptions in Australian Parliament by Gender",
    x = "Year",
    y = "Total Interruptions",
    color = "Gender"
  ) +
  scale_color_brewer(palette = "Set2") +  # Use a colorblind-friendly palette
  theme_minimal(base_size = 14) +  # Adjust base font size for readability
  theme(
    legend.position = "top",  # Move legend to the top
    legend.title = element_text(size = 10, face = "bold"),  # Bold legend title
    legend.text = element_text(size = 8),  # Adjust legend text size
    plot.title = element_text(size = 14, face = "bold"),  # Bold title
    axis.title = element_text(size = 12),  # Axis title size
    axis.text = element_text(size = 11)  # Axis text size
  ) +
  scale_x_continuous(breaks = seq(min(final_data$year), max(final_data$year), by = 5))  # Adjust x-axis breaks

```

```{r}
#| include: false
#| warning: false
#| message: false

# Ensure that 'date' is correctly formatted as Date and 'year' is extracted
hansard_corpus <- hansard_corpus %>%
  mutate(
    date = as.Date(date, format = "%Y-%m-%d"),  # Convert date if it's in character form
    year = if_else(is.na(date), NA_integer_, year(date))  # Extract year, handle NA dates
  )

# Assuming hansard_corpus has the year extracted already. If not, add this in the mutate step.
interruptions_by_year_gender <- hansard_corpus %>%
  group_by(year, gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    interruption_proportion = total_interruptions / total_speeches
  )

# Calculate for all genders per year
interruptions_by_year_all <- hansard_corpus %>%
  group_by(year) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    gender = "All",
    interruption_proportion = total_interruptions / total_speeches
  )

# Combine data
final_data <- bind_rows(interruptions_by_year_gender, interruptions_by_year_all)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-2
#| fig-cap: "Yearly Interruptions Proportions in Australian Parliament by Gender"

ggplot(final_data, aes(x = year, y = interruption_proportion, color = gender)) +
  geom_line(size = 1.2) +  # Thicker lines for better visibility
  geom_point(size = 2) +  # Larger points to highlight data
  labs(
    title = "Yearly Interruptions Proportions\nin Australian Parliament by Gender",  # Add \n for a line break
    x = "Year",
    y = "Total Interruptions Proportions",
    color = "Gender"
  ) +
  scale_color_brewer(palette = "Set2") +  # Use a colorblind-friendly palette
  theme_minimal(base_size = 14) +  # Adjust base font size for readability
  theme(
    legend.position = "top",  # Move legend to the top
    legend.title = element_text(size = 10, face = "bold"),  # Bold legend title
    legend.text = element_text(size = 8),  # Adjust legend text size
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center the title
    axis.title = element_text(size = 12),  # Axis title size
    axis.text = element_text(size = 11)  # Axis text size
  ) +
  scale_x_continuous(breaks = seq(min(final_data$year), max(final_data$year), by = 5))  # Adjust x-axis breaks

```

@fig-1 illustrates interruptions in the Australian Parliament from 1998 to 2022, categorized by gender. Notably, male speakers consistently faced more interruptions than females, with a peak in 2011 and a subsequent decline. Female speakers experienced fewer interruptions, with relatively stable figures across the years. The "Unknown" category, consistently low, suggests either minimal occurrence or documentation of such interruptions.

To provide a fair comparison of interruptions across different genders within the Australian Parliament, we normalize the data by calculating proportions. This method addresses the gender imbalance in parliament, ensuring that the analysis reflects the frequency of interruptions relative to each gender's representation. @fig-2 presents the proportions of speeches interrupted by gender. It indicates that male speakers had a higher proportion of their speeches interrupted, especially during the mid-2000s to early 2010s, though the gap between genders has narrowed over time. The proportion of interruptions for female speakers generally mirrors these fluctuations, implying that specific topics or periods may have incited more active parliamentary interactions. The "Unknown" gender shows minimal changes, highlighting its limited data representation. This analysis points out both the consistency and evolution of gender dynamics within parliamentary debates over the years.

```{r}
#| include: false
#| warning: false
#| message: false
# Summarize interruptions by year, party, and gender
interruptions_summary <- hansard_corpus %>%
  group_by(year, party, gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    .groups = 'drop'
  )

```

```{r}
#| include: false
#| warning: false
#| message: false
# Plotting the data as a grouped bar chart
ggplot(interruptions_summary, aes(x = party, y = total_interruptions, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge()) +  # Use position_dodge() for grouped bars
  labs(
    title = "Interruptions by Gender and Party",
    x = "Party",
    y = "Total Interruptions",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability

```

```{r}
#| include: false
#| warning: false
#| message: false
# Filter data for only LP and ALP parties and summarize interruptions by gender
interruptions_major_parties <- hansard_corpus %>%
  filter(party %in% c("LP", "ALP")) %>%
  mutate(
    interject = as.numeric(interject)  # Ensure interject is numeric
  ) %>%
  group_by(party, gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    .groups = 'drop'
  )

```

```{r}
#| include: false
#| warning: false
#| message: false
# Assuming the 'hansard_corpus' and 'interject' are properly loaded and converted
# First, let's create a summary for each gender within the two parties.
interruptions_gender <- hansard_corpus %>%
  filter(party %in% c("LP", "ALP")) %>%
  group_by(party, gender) %>%
  summarise(
    total_interruptions = sum(as.numeric(interject), na.rm = TRUE),
    .groups = 'drop'
  )

```

```{r}
#| include: false
#| warning: false
#| message: false
# Plotting with the inclusion of "All"
ggplot(interruptions_major_parties, aes(x = party, y = total_interruptions, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +
  scale_fill_manual(values = c("Female" = "#FF9999", "Male" = "#9999FF")) +  # Color for "All"
  labs(
    title = "Gender-based and Total Interruptions\nin Major Australian Political Parties",
    x = "Political Party",
    y = "Total Interruptions",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust=0.5),
    plot.subtitle = element_text(face = "italic"),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(angle = 0, vjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )

```
```{r}
#| include: false
#| warning: false
#| message: false
# Calculate the normalized proportion (probability of interruption)
interruptions_normalized <- hansard_corpus %>%
  filter(party %in% c("LP", "ALP")) %>%
  group_by(party, gender) %>%
  summarise(
    total_interruptions = sum(as.numeric(interject), na.rm = TRUE),  # Total interruptions by gender and party
    total_speeches = n(),  # Total speeches by gender and party
    .groups = 'drop'
  ) %>%
  mutate(
    interruption_probability = total_interruptions / total_speeches  # Normalized proportion (probability)
  )

```

```{r}
#| include: false
#| warning: false
#| message: false
# Plotting the normalized proportions
ggplot(interruptions_normalized, aes(x = party, y = interruption_probability, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +  # Bar plot with proportions
  scale_fill_manual(values = c("Female" = "#FF9999", "Male" = "#9999FF")) +  # Custom colors for gender
  labs(
    title = "Normalized Gender-based Interruptions\nin Major Australian Political Parties",
    x = "Political Party",
    y = "Probability of Interruption",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(angle = 0, vjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-3
#| fig-cap: "Interruptions in Major Australian Political Parties by Gender"

# Remove the legend from the first plot
plot1 <- ggplot(interruptions_major_parties, aes(x = party, y = total_interruptions, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +
  scale_fill_manual(values = c("Female" = "#FF9999", "Male" = "#9999FF")) +  # Colors for genders
  labs(
    x = "Political Party",
    y = "Total Interruptions"
  ) +
  theme_minimal() +
  theme(
    plot.subtitle = element_text(face = "bold", size = 12, hjust = 0.5),  # Subtitle styling
    axis.title = element_text(size = 11),
    axis.text.x = element_text(angle = 0, vjust = 0.5),
    legend.position = "none",  # Remove the legend
    legend.text = element_text(size = 9)
  )

# Keep the legend only in the second plot
plot2 <- ggplot(interruptions_normalized, aes(x = party, y = interruption_probability, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +  # Bar plot with proportions
  scale_fill_manual(values = c("Female" = "#FF9999", "Male" = "#9999FF")) +  # Custom colors for gender
  labs(
    x = "Political Party",
    y = "Probability of Interruption",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.subtitle = element_text(face = "bold", size = 12, hjust = 0.5),  # Subtitle styling
    axis.title = element_text(size = 11),
    axis.text.x = element_text(angle = 0, vjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )

# Add a combined title and arrange plots with patchwork
(plot1 + plot2) +
  plot_annotation(
    title = "Interruptions in Major Australian Political Parties by Gender",
    theme = theme(
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)  # Big title styling
    )
  )

```

@fig-3 focuses on the Liberal Party (LP) and the Australian Labor Party (ALP), the two major parties in Australian politics, to analyze interruptions by gender. Since these parties dominate parliamentary proceedings, they provide a significant sample of gender interaction dynamics. The total interruptions are higher in the ALP compared to the LP, with male members experiencing more interruptions than female members in both parties. We use proportions to normalize the data, accounting for differences in gender representation within these parties. This approach helps us understand the rate at which female and male members are interrupted relative to their participation, offering a clearer insight into the parliamentary discourse dynamics.

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
git config --global http.postBuffer 524288000

git config --global http.postBuffer 629145600

git reset --soft HEAD~1

git lfs install
git lfs track "*.parquet"
git add .gitattributes
git commit -m "Track large parquet files with Git LFS"
git add data/01-raw_data/hansard-corpus/hansard_corpus_1998_to_2022.parquet
git commit -m "Add large Hansard Parquet file"
git push origin main

# Data {#sec-data}

## Overview

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

\newpage

# References