### 2025-01-10 paper/paper.qmd
add in.gov, count, etc.

---
title: "Voices in Parliament: Analyzing Interruptions Across Gender in the Australian Political Landscape"
subtitle: "My subtitle if needed"
author: 
  - Hannah Yu
  - Rohan Alexander
thanks: "Code and data are available at: https://github.com/hannahyu07/Hansard_Interruptions"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction {#sec-introduction}

Although interjections are officially prohibited during a member's speech in the Australian Parliament, interruptions still occur frequently, particularly during heated debates often fueled by political disagreements [@aphInterruptions2025]. This raises crucial questions about gender dynamics within the context of parliamentary debates in Australia. Concerns are growing regarding whether women are disproportionately interrupted compared to men, highlighting potential underlying biases that could influence the effectiveness and fairness of parliamentary proceedings.

There has been considerable research on speech patterns and gender-based interruptions in government speeches across various countries. For instance, during U.S. Supreme Court confirmation hearings, female and minority nominees are interrupted more frequently than their male and white counterparts [@genderRaceSCCHearings2024]. Additionally, in the US Congress, women also experienced more interruptions than men, especially during discussions of women’s issues, where women are more than twice as likely to be interrupted compared to non-gendered topics [@genderInterruptionsCongress2022]. In the Ecuadorian Congress, female legislators not only deliver fewer speeches but also speak for shorter durations compared to their male peers. Furthermore, women often wait longer to regain the floor after interruptions, leading them to strategically shorten their speeches to minimize further disruptions. However, women in authoritative positions can mitigate some of these interruption-related penalties, with senior women's speeches often exceeding those of men in length when they achieve comparable levels of experience or hold significant roles [@vera2021politics].

Contrastingly, studies of the Australian High Court reveal a different dynamic [@jacobiPredictabilityInterruptions]. The result suggested that contrary to the U.S. Supreme Court, there is no significant difference between female and male justices regarding interruptions. Although women experience a higher rate of interruptions relative to their speaking time, the subtlety of this effect suggests that gender dynamics within the Australian court might be more balanced compared to the U.S. Supreme Court. This highlights the differences in how judicial interactions are influenced by gender across different legal systems.

Building on existing literature suggesting that the Australian High Court may exhibit better gender dynamics than its counterparts in the US, we aim to delve deeper into the issue of gender-based interruptions within the Australian Parliament. Utilizing the Australian Hansard, the official written record of parliamentary debates which has been digitized by Lindsay Katz and Rohan Alexander, we have access to comprehensive records of all parliamentary proceedings from 1998 to 2022 [@katz2023digitization]. From the dataset, we aim to analyze the frequency and context of interruptions based on gender. Specifically, our study will focus on comparing the rates at which male and female parliamentarians are interrupted during their speeches.

The remainder of this paper is structured as follows: @sec-data demonstrates the data used for this paper, @sec-model describes, justifies, and explains my model, @sec-results highlights the results of the predictions using tables and graphs, and @sec-discussion contains important implications and conclusions based on the findings.

# Data {#sec-data}

# Model {#sec-model}
# Results {#sec-results}

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(here)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(knitr)
library(tibble)
library(gt)
library(patchwork)
library(kableExtra)
library(dplyr)
library(stringr)

```

```{r}
#| include: false
#| warning: false
#| message: false
hansard_corpus <- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))
```

```{r}
#| include: false
#| warning: false
#| message: false
# View the first few rows
head(hansard_corpus, 150)

```

```{r}
#| include: false
#| warning: false
#| message: false
summary(hansard_corpus)

```

```{r}
#| include: false
#| warning: false
#| message: false
# Calculate interruptions and speeches by gender, including rounding the proportion
interruptions_by_gender <- hansard_corpus %>%
  group_by(gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    .groups = 'drop'  # Automatically drop the grouping
  ) %>%
  mutate(
    interruption_proportion = round(total_interruptions / total_speeches, 2)  # Calculate and round the proportion of interruptions
  )

# Calculate total interruptions and speeches for all genders combined, with rounding
interruptions_all <- hansard_corpus %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n()
  ) %>%
  mutate(
    gender = "All",
    interruption_proportion = round(total_interruptions / total_speeches, 2)  # Round the proportion here as well
  )

# Combine the results for specific genders and "All"
final_results <- bind_rows(interruptions_by_gender, interruptions_all)

# Print the results
print(final_results)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Gender-Based Interruption Analysis in Parliamentary Speeches "
#| label: tbl-1

# Create a clean and professional table
final_results %>%
  gt() %>%
  tab_header(
    title = "Proportion of Interruptions by Gender",
  ) %>%
  fmt_percent(
    columns = interruption_proportion,
    decimals = 2
  ) %>%
  cols_label(
    gender = "Gender",
    total_interruptions = "Total Interruptions",
    total_speeches = "Total Speeches",
    interruption_proportion = "Interruption Proportion"
  ) %>%
  tab_style(
    style = cell_borders(sides = "all", color = "gray", weight = px(1)),
    locations = cells_body()
  ) %>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(18),
    heading.subtitle.font.size = px(14),
    data_row.padding = px(5),
    table.border.top.color = "gray",
    table.border.bottom.color = "gray"
  )%>%
  tab_style(
    style = cell_text(align = "left"),  # Align text to the left
    locations = cells_body(columns = everything())  # Apply to all columns in the body
  )

```

@tbl-1 provides a clear view of the proportion of interruptions by gender across all speeches in the dataset. The results show that male speakers were interrupted more frequently (21% of speeches) compared to female speakers (18% of speeches). Interestingly, the interruption proportion for unknown gender speakers is significantly lower, at only 11%. Overall, considering all speeches regardless of gender, the interruption proportion stands at 18%. This data suggests that gender may play a role in how frequently speakers are interrupted during parliamentary sessions, with male speakers facing a higher rate of interruption. This could reflect broader dynamics of gender interaction within the parliamentary setting.

```{r}
#| include: false
#| warning: false
#| message: false
#Based on the following data it shows that in fact men seem to be interrupted more on avergae than women in the australian parliment, but this could be changed from the "unknown" category.
# Summarize interruptions by year and gender
interruptions_summary <- hansard_corpus %>%
  group_by(year, gender) %>%
  summarise(
    total_interruptions = sum(as.numeric(interject), na.rm = TRUE),
    .groups = 'drop'
  )

# Summarize total interruptions for all genders combined per year
interruptions_all <- hansard_corpus %>%
  group_by(year) %>%
  summarise(
    total_interruptions = sum(as.numeric(interject), na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    gender = "All"
  )

# Combine gender-specific and "All" categories data
final_data <- bind_rows(interruptions_summary, interruptions_all)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-1
#| fig-cap: "Yearly Interruptions in Australian Parliament by Gender"

ggplot(final_data, aes(x = year, y = total_interruptions, color = gender)) +
  geom_line(size = 1.2) +  # Thicker lines for better visibility
  geom_point(size = 2) +  # Larger points to highlight data
  labs(
    title = "Yearly Interruptions in Australian\nParliament by Gender",
    x = "Year",
    y = "Total Interruptions",
    color = "Gender"
  ) +
  scale_color_brewer(palette = "Set2") +  # Use a colorblind-friendly palette
  theme_minimal(base_size = 14) +  # Adjust base font size for readability
  theme(
    legend.position = "top",  # Move legend to the top
    legend.title = element_text(size = 10, face = "bold"),  # Bold legend title
    legend.text = element_text(size = 8),  # Adjust legend text size
    plot.title = element_text(size = 14, face = "bold",  hjust = 0.5),  # Bold title
    axis.title = element_text(size = 12),  # Axis title size
    axis.text = element_text(size = 11)  # Axis text size
  ) +
  scale_x_continuous(breaks = seq(min(final_data$year), max(final_data$year), by = 5))  # Adjust x-axis breaks

```

```{r}
#| include: false
#| warning: false
#| message: false

# Ensure that 'date' is correctly formatted as Date and 'year' is extracted
hansard_corpus <- hansard_corpus %>%
  mutate(
    date = as.Date(date, format = "%Y-%m-%d"),  # Convert date if it's in character form
    year = if_else(is.na(date), NA_integer_, year(date))  # Extract year, handle NA dates
  )

# Assuming hansard_corpus has the year extracted already. If not, add this in the mutate step.
interruptions_by_year_gender <- hansard_corpus %>%
  group_by(year, gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    interruption_proportion = total_interruptions / total_speeches
  )

# Calculate for all genders per year
interruptions_by_year_all <- hansard_corpus %>%
  group_by(year) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    gender = "All",
    interruption_proportion = total_interruptions / total_speeches
  )

# Combine data
final_data1 <- bind_rows(interruptions_by_year_gender, interruptions_by_year_all)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-2
#| fig-cap: "Yearly Interruptions Proportions in Australian Parliament by Gender"

ggplot(final_data1, aes(x = year, y = interruption_proportion, color = gender)) +
  geom_line(size = 1.2) +  # Thicker lines for better visibility
  geom_point(size = 2) +  # Larger points to highlight data
  labs(
    title = "Yearly Interruptions Proportions\nin Australian Parliament by Gender",  # Add \n for a line break
    x = "Year",
    y = "Total Interruptions Proportions",
    color = "Gender"
  ) +
  scale_color_brewer(palette = "Set2") +  # Use a colorblind-friendly palette
  theme_minimal(base_size = 14) +  # Adjust base font size for readability
  theme(
    legend.position = "top",  # Move legend to the top
    legend.title = element_text(size = 10, face = "bold"),  # Bold legend title
    legend.text = element_text(size = 8),  # Adjust legend text size
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center the title
    axis.title = element_text(size = 12),  # Axis title size
    axis.text = element_text(size = 11)  # Axis text size
  ) +
  scale_x_continuous(breaks = seq(min(final_data$year), max(final_data$year), by = 5))  # Adjust x-axis breaks

```

@fig-1 illustrates interruptions in the Australian Parliament from 1998 to 2022, categorized by gender. Notably, male speakers consistently faced more interruptions than females, with a peak in 2011 and a subsequent decline. Female speakers experienced fewer interruptions, with relatively stable figures across the years. The "Unknown" category, consistently low, suggests either minimal occurrence or documentation of such interruptions.

To provide a fair comparison of interruptions across different genders within the Australian Parliament, we normalize the data by calculating proportions. This method addresses the gender imbalance in parliament, ensuring that the analysis reflects the frequency of interruptions relative to each gender's representation. @fig-2 presents the proportions of speeches interrupted by gender. It indicates that male speakers had a higher proportion of their speeches interrupted, especially during the mid-2000s to early 2010s, though the gap between genders has narrowed over time. The proportion of interruptions for female speakers generally mirrors these fluctuations, implying that specific topics or periods may have incited more active parliamentary interactions. The "Unknown" gender shows minimal changes, highlighting its limited data representation. This analysis points out both the consistency and evolution of gender dynamics within parliamentary debates over the years.

```{r}
#| include: false
#| warning: false
#| message: false
print(final_data1)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-2_chamber
#| fig-cap: "Yearly Interruptions Proportions in Australian Parliament by Gender (Chamber)"

no_fedchamb <- hansard_corpus %>%
  filter(fedchamb_flag == 0)

# Group by year and gender to summarize interruptions
interruptions_by_year_gender <- no_fedchamb %>%
  group_by(year, gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    interruption_proportion = total_interruptions / total_speeches
  )

# Calculate interruptions for all genders combined per year
interruptions_by_year_all <- no_fedchamb %>%
  group_by(year) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    gender = "All",
    interruption_proportion = total_interruptions / total_speeches
  )

# Combine the gender-specific and overall data
final_data2 <- bind_rows(interruptions_by_year_gender, interruptions_by_year_all)

# Plotting the interruptions proportions by year and gender
ggplot(final_data2, aes(x = year, y = interruption_proportion, color = gender)) +
  geom_line(size = 1.2) +  # Use thicker lines for better visibility
  geom_point(size = 2) +  # Use larger points to highlight data points
  labs(
    title = "Yearly Interruptions Proportions\nin Australian Parliament by Gender (Chamber)",
    x = "Year",
    y = "Total Interruptions Proportions",
    color = "Gender"
  ) +
  scale_color_brewer(palette = "Set2") +  # Use a colorblind-friendly palette
  theme_minimal(base_size = 14) +  # Use a minimal theme with readable base font size
  theme(
    legend.position = "top",  # Position legend at the top for better accessibility
    legend.title = element_text(size = 10, face = "bold"),  # Bold the legend title
    legend.text = element_text(size = 8),  # Set legend text size
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center and bold the plot title
    axis.title = element_text(size = 12),  # Set axis title size
    axis.text = element_text(size = 11)  # Set axis text size
  ) +
  scale_x_continuous(breaks = seq(min(final_data$year), max(final_data$year), by = 5))  # Set x-axis breaks every 5 years

```
```{r}
#| include: false
#| warning: false
#| message: false
# Displaying the table of final data
print(final_data2)

```

```{r}
#| include: false
#| warning: false
#| message: false
# Pivot to get male and female data on the same row
comparison_data <- final_data2 %>%
  pivot_wider(
    id_cols = year,
    names_from = gender,
    values_from = interruption_proportion
  )

# Determine which gender had more interruptions per year
yearly_comparison <- comparison_data %>%
  mutate(
    More_Interrupted_Gender = case_when(
      Male > Female ~ "Male",
      Female > Male ~ "Female",
      TRUE ~ "Equal"
    )
  )

# Count the number of years each gender was interrupted more
gender_counts <- yearly_comparison %>%
  count(More_Interrupted_Gender)

```

```{r}
#| include: false
#| warning: false
#| message: false
# Determine which gender had more interruptions per year
yearly_comparison <- comparison_data %>%
  mutate(
    More_Interrupted_Gender = case_when(
      Male > Female ~ "Male",
      Female > Male ~ "Female",
      TRUE ~ "Equal"  # This handles cases where the interruption proportions are the same
    )
  ) %>%
  select(year, More_Interrupted_Gender)  # Select only the year and the more interrupted gender columns

# Print the simplified table
print(yearly_comparison)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Gender Interruption Proportions Annual Comparison Across General and Chamber"
#| label: tbl-compareyear
# Prepare the data
comparison_all <- final_data1 %>%
  pivot_wider(
    id_cols = year,
    names_from = gender,
    values_from = interruption_proportion
  ) %>%
  mutate(
    More_Interrupted_Gender_All = case_when(
      Male > Female ~ "Male",
      Female > Male ~ "Female",
      TRUE ~ "Equal"
    )
  )

comparison_chamber <- final_data2 %>%
  pivot_wider(
    id_cols = year,
    names_from = gender,
    values_from = interruption_proportion
  ) %>%
  mutate(
    More_Interrupted_Gender_Chamber = case_when(
      Male > Female ~ "Male",
      Female > Male ~ "Female",
      TRUE ~ "Equal"
    )
  )

# Assuming final_comparison is prepared as follows
final_comparison <- comparison_all %>%
  select(year, More_Interrupted_Gender_All) %>%
  left_join(comparison_chamber %>% select(year, More_Interrupted_Gender_Chamber), by = "year")

final_comparison %>%
  gt %>%
  cols_label(
    year = "Year",
    More_Interrupted_Gender_All = "More Interrupted Gender All",
    More_Interrupted_Gender_Chamber = "More Interrupted Gender Chamber")%>%
  tab_style(
    style = cell_text(align = "left"),  # Align text to the left
    locations = cells_body(columns = everything())  # Apply to all columns in the body
  )

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-compare
#| fig-cap: "Yearly Interruptions Proportions in Australian Parliament by Gender"

# Original dataset, no condition on fedchamb_flag
interruptions_original <- hansard_corpus %>%
  filter(gender %in% c("Male", "Female")) %>%
  mutate(
    date = as.Date(date, format = "%Y-%m-%d"),
    year = if_else(is.na(date), NA_integer_, year(date))
  ) %>%
  group_by(year, gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    interruption_proportion = total_interruptions / total_speeches,
    condition = "All Data"
  )

# Dataset with fedchamb_flag == 0
interruptions_fedchamb_flag_0 <- hansard_corpus %>%
  filter(gender %in% c("Male", "Female"), fedchamb_flag == 0) %>%
  mutate(
    date = as.Date(date, format = "%Y-%m-%d"),
    year = if_else(is.na(date), NA_integer_, year(date))
  ) %>%
  group_by(year, gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    interruption_proportion = total_interruptions / total_speeches,
    condition = "fedchamb_flag = 0"
  )

# Combine the datasets
final_data3 <- bind_rows(interruptions_original, interruptions_fedchamb_flag_0)

# Adjust labels for condition and linetype
ggplot(final_data3, aes(x = year, y = interruption_proportion, color = gender, linetype = condition)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "Yearly Interruptions Proportions in\nAustralian Parliament by Gender",
    subtitle = "Comparing All Data vs. Chamber Data Only",
    x = "Year",
    y = "Interruption Proportion",
    color = "Gender",
    linetype = "Dataset Type"
  ) +
  scale_color_brewer(palette = "Set2") +
  scale_linetype_manual(values = c("solid", "dotted"), 
                        labels = c("All", "Chamber")) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    legend.title = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 8),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, face = "italic", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11)
  ) +
  scale_x_continuous(breaks = seq(min(final_data3$year), max(final_data3$year), by = 5))  # Adjust x-axis breaks

```

Only analyzing Chamber members, @fig-2_chamber again illustrates interruptions in the Australian Parliament from 1998 to 2022, categorized by gender. To distinguish the differences across the two figures, @fig-compare contains both the interruptions for all data and Chamber member data. While the interruption patterns in the chamber-only data in @fig-2_chamber largely mirror those observed in the comprehensive dataset @fig-2, there are slight increases in the interruption proportions for women, indicating a potentially heightened level of disruption experienced by female members in non-federal chambers.

@tbl-compareyear displays the more interrupted gender every year in the parliament and compares all members and only chamber members. While most years are the same, there is a notable shift in the gender experiencing more frequent interruptions from 2018 to 2020. In the broader dataset, males were more frequently interrupted; however, this pattern shifts in the non-federal chamber data, where females became the more interrupted gender during these years. This change highlights the variability in gender dynamics within different parliamentary environments and suggests that factors specific to non-federal chamber settings may influence these interaction patterns more significantly for female members.

```{r}
#| include: false
#| warning: false
#| message: false
# Filter and aggregate data based on the 'question' flag and 'gender'
interruptions_question_analysis <- hansard_corpus %>%
  filter(gender %in% c("Male", "Female")) %>%
  group_by(gender, is_question = as.factor(question)) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    total_speeches = sum(total_speeches),
    interruption_proportion = total_interruptions / total_speeches,
    .groups = 'drop'
  )

# Print the resulting dataset to verify
print(interruptions_question_analysis)

```

```{r}
#| include: false
#| warning: false
#| message: false
# Filter and aggregate data based on the 'question' flag and 'gender'
interruptions_question_analysis <- hansard_corpus %>%
  filter(gender %in% c("Male", "Female")) %>%
  group_by(gender, is_answer = as.factor(answer)) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    total_speeches = sum(total_speeches),
    interruption_proportion = total_interruptions / total_speeches,
    .groups = 'drop'
  )

# Print the resulting dataset to verify
print(interruptions_question_analysis)
```

```{r}
#| include: false
#| warning: false
#| message: false
# Summarize interruptions by year, party, and gender
interruptions_summary <- hansard_corpus %>%
  group_by(year, party, gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    .groups = 'drop'
  )

```

```{r}
#| include: false
#| warning: false
#| message: false
# Plotting the data as a grouped bar chart
ggplot(interruptions_summary, aes(x = party, y = total_interruptions, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge()) +  # Use position_dodge() for grouped bars
  labs(
    title = "Interruptions by Gender and Party",
    x = "Party",
    y = "Total Interruptions",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability

```

```{r}
#| include: false
#| warning: false
#| message: false
# Filter data for only LP and ALP parties and summarize interruptions by gender
interruptions_major_parties <- hansard_corpus %>%
  filter(party %in% c("LP", "ALP")) %>%
  mutate(
    interject = as.numeric(interject)  # Ensure interject is numeric
  ) %>%
  group_by(party, gender) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    .groups = 'drop'
  )

```

```{r}
#| include: false
#| warning: false
#| message: false
# Assuming the 'hansard_corpus' and 'interject' are properly loaded and converted
# First, let's create a summary for each gender within the two parties.
interruptions_gender <- hansard_corpus %>%
  filter(party %in% c("LP", "ALP")) %>%
  group_by(party, gender) %>%
  summarise(
    total_interruptions = sum(as.numeric(interject), na.rm = TRUE),
    .groups = 'drop'
  )

```

```{r}
#| include: false
#| warning: false
#| message: false
# Plotting with the inclusion of "All"
ggplot(interruptions_major_parties, aes(x = party, y = total_interruptions, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +
  scale_fill_manual(values = c("Female" = "#FF9999", "Male" = "#9999FF")) +  # Color for "All"
  labs(
    title = "Gender-based and Total Interruptions\nin Major Australian Political Parties",
    x = "Political Party",
    y = "Total Interruptions",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust=0.5),
    plot.subtitle = element_text(face = "italic"),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(angle = 0, vjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )

```

```{r}
#| include: false
#| warning: false
#| message: false
# Calculate the normalized proportion (probability of interruption)
interruptions_normalized <- hansard_corpus %>%
  filter(party %in% c("LP", "ALP")) %>%
  group_by(party, gender) %>%
  summarise(
    total_interruptions = sum(as.numeric(interject), na.rm = TRUE),  # Total interruptions by gender and party
    total_speeches = n(),  # Total speeches by gender and party
    .groups = 'drop'
  ) %>%
  mutate(
    interruption_probability = total_interruptions / total_speeches  # Normalized proportion (probability)
  )

```

```{r}
#| include: false
#| warning: false
#| message: false
# Plotting the normalized proportions
ggplot(interruptions_normalized, aes(x = party, y = interruption_probability, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +  # Bar plot with proportions
  scale_fill_manual(values = c("Female" = "#FF9999", "Male" = "#9999FF")) +  # Custom colors for gender
  labs(
    title = "Normalized Gender-based Interruptions\nin Major Australian Political Parties",
    x = "Political Party",
    y = "Probability of Interruption",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(angle = 0, vjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-3
#| fig-cap: "Interruptions in Major Australian Political Parties by Gender"

# Remove the legend from the first plot
plot1 <- ggplot(interruptions_major_parties, aes(x = party, y = total_interruptions, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +
  scale_fill_manual(values = c("Female" = "#FF9999", "Male" = "#9999FF")) +  # Colors for genders
  labs(
    x = "Political Party",
    y = "Total Interruptions"
  ) +
  theme_minimal() +
  theme(
    plot.subtitle = element_text(face = "bold", size = 12, hjust = 0.5),  # Subtitle styling
    axis.title = element_text(size = 11),
    axis.text.x = element_text(angle = 0, vjust = 0.5),
    legend.position = "none",  # Remove the legend
    legend.text = element_text(size = 9)
  )

# Keep the legend only in the second plot
plot2 <- ggplot(interruptions_normalized, aes(x = party, y = interruption_probability, fill = gender)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +  # Bar plot with proportions
  scale_fill_manual(values = c("Female" = "#FF9999", "Male" = "#9999FF")) +  # Custom colors for gender
  labs(
    x = "Political Party",
    y = "Probability of Interruption",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.subtitle = element_text(face = "bold", size = 12, hjust = 0.5),  # Subtitle styling
    axis.title = element_text(size = 11),
    axis.text.x = element_text(angle = 0, vjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )

# Add a combined title and arrange plots with patchwork
(plot1 + plot2) +
  plot_annotation(
    title = "Interruptions in Major Australian Political Parties by Gender",
    theme = theme(
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)  # Big title styling
    )
  )

```

@fig-3 focuses on the Liberal Party (LP) and the Australian Labor Party (ALP), the two major parties in Australian politics, to analyze interruptions by gender. Since these parties dominate parliamentary proceedings, they provide a significant sample of gender interaction dynamics. The total interruptions are higher in the ALP compared to the LP, with male members experiencing more interruptions than female members in both parties. We use proportions to normalize the data, accounting for differences in gender representation within these parties. This approach helps us understand the rate at which female and male members are interrupted relative to their participation, offering a clearer insight into the parliamentary discourse dynamics.

```{r}
#| include: false
#| warning: false
#| message: false
# Calculate basic summary statistics for in.gov
in_gov_stats <- hansard_corpus %>%
  summarise(
    Count_In_Gov = sum(in.gov == 1, na.rm = TRUE),
    Count_Not_In_Gov = sum(in.gov == 0, na.rm = TRUE),
    Count_NA = sum(is.na(in.gov)),
    Total = n()
  )

print(in_gov_stats)

```

```{r}
#| include: false
#| warning: false
#| message: false
# Calculate basic summary statistics for in.gov
interject_stats <- hansard_corpus %>%
  summarise(
    Count_Interject = sum(interject == 1, na.rm = TRUE),
    Count_No_Interject = sum(interject == 0, na.rm = TRUE),
    Count_NA = sum(is.na(interject)),
    Total = n()
  )

print(interject_stats)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Interruption Summary by Government Status (All)"
#| label: tbl-ingov1

# Assume that interruption_summary is already prepared as previously coded
interruption_summary <- hansard_corpus %>%
  mutate(
    in_gov = factor(in.gov, levels = c(0, 1), labels = c("In Opposition", "In Government")),
    interject = factor(interject, levels = c(0, 1), labels = c("No", "Yes"))
  ) %>%
  group_by(in_gov, interject) %>%
  summarise(
    Count = n(),
    .groups = 'drop'
  ) %>%
  arrange(in_gov, desc(interject)) %>%
  rename(
    `Government Status` = in_gov,
    `Interruption` = interject,
    `Count` = Count
  )

# Create a clean and professional table
interruption_summary %>%
  gt() %>%
  tab_style(
    style = cell_text(align = "left"),  # Align text to the left
    locations = cells_body(columns = everything())  # Apply to all columns in the body
  )

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Interruption Summary by Government Status in Non-Federal Chamber Sessions"
#| label: tbl-ingov2

# Assume that interruption_summary is already prepared as previously coded
interruption_summary <- hansard_corpus %>%
  filter(fedchamb_flag == 0) %>%
  mutate(
    in_gov = factor(in.gov, levels = c(0, 1), labels = c("In Opposition", "In Government")),
    interject = factor(interject, levels = c(0, 1), labels = c("No", "Yes"))
  ) %>%
  group_by(in_gov, interject) %>%
  summarise(
    Count = n(),
    .groups = 'drop'
  ) %>%
  arrange(in_gov, desc(interject)) %>%
  rename(
    `Government Status` = in_gov,
    `Interruption` = interject,
    `Count` = Count
  )
interruption_summary %>%
  gt() %>%
  tab_style(
    style = cell_text(align = "left"),  # Align text to the left
    locations = cells_body(columns = everything())  # Apply to all columns in the body
  )

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-gov
#| fig-cap: "Count of Interruptions by Government Status and Gender"

# Prepare the data with absolute counts
interruption_counts <- hansard_corpus %>%
  filter(gender %in% c("Male", "Female")) %>%
  filter(fedchamb_flag == 0) %>%
  mutate(
    in_gov = factor(ifelse(is.na(in.gov), "NA", ifelse(in.gov == 1, "In Government", "In Opposition")),
                    levels = c("In Opposition", "In Government", "NA")),
    interject = factor(interject, levels = c(0, 1), labels = c("No", "Yes"))
  ) %>%
  group_by(in_gov, interject, gender) %>%
  summarise(
    Count = n(),
    .groups = 'drop'
  )

# Plot using ggplot2 with updated colors and adjusted x-axis labels
ggplot(interruption_counts, aes(x = in_gov, y = Count, fill = interject)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ gender, scales = "free_y") +
  labs(
    title = "Interruptions Count by Government Status and Gender",
    x = "Government Status",
    y = "Count of Interruptions",
    fill = "Interruption Occurred"
  ) +
  scale_fill_manual(values = c("No" = "steelblue", "Yes" = "salmon")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 0, vjust = 2, hjust = 0.5), # Adjust label positioning
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.position = "bottom"
  )

```

@tbl-ingov1, @tbl-ingov2, and @fig-gov offer a detailed look at interruptions during parliamentary sessions, categorized by members' government status and gender. Notably, interruptions are extremely rare for members with clearly identified status. This indicates a potential correlation between established government status and more orderly interactions, highlighting the importance of role clarity in maintaining decorum.

However, the significant amount of missing data makes reliable inferences about the true nature of interruptions and their correlation with member status challenging. The vast majority of interruptions occur when the government status of the speaker is unknown. This trend is consistent across both the comprehensive dataset and the subset limited to non-federal chamber sessions. Interruptions are exceedingly rare for members whose government status is known, with only 22 interruptions for those "In Government" and an even fewer 4 for those "In Opposition" across all data. Due to this substantial amount of missing information, it is very difficult to firmly conclude whether our observed patterns hold true.

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Summary of Parliamentary Session Activities and Interruptions"
#| label: tbl-qa_interrupt1
# Assuming hansard_corpus is already loaded and prepped
summary_stats <- hansard_corpus %>%
  mutate(
    question_numeric = as.numeric(as.character(question)),
    answer_numeric = as.numeric(as.character(answer))
  ) %>%
  summarise(
    Questions = sum(question_numeric, na.rm = TRUE),
    Answers = sum(answer_numeric, na.rm = TRUE),
    Total_Interruptions = sum(interject, na.rm = TRUE),
    Interruptions_During_Questions = sum(interject[question_numeric == 1], na.rm = TRUE),
    Interruptions_During_Answers = sum(interject[answer_numeric == 1], na.rm = TRUE)
  )

summary_stats %>%
  gt() %>%
  cols_label(
    Questions = "Questions",
    Answers = "Answers",
    Total_Interruptions = "Total Interruptions",
    Interruptions_During_Questions = "Question Interrupted",
    Interruptions_During_Answers = "Answer Interrupted"
  ) %>%
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_body(columns = everything())
  ) %>%
  tab_options(
    table.width = pct(100),  # Set the table width to 100% of the container
    table.align = "left"  # Align the table to the left of the container
  ) %>%
  tab_style(
    style = cell_text(align = "left"),  # Align text to the left
    locations = cells_body(columns = everything())  # Apply to all columns in the body
  )

```

```{r}
#| include: false
#| warning: false
#| message: false
# Review entries where questions or answers are marked
question_instances <- hansard_corpus %>%
  filter(question == 1)
print(question_instances)
```

```{r}
#| include: false
#| warning: false
#| message: false
# Filter and aggregate data based on the 'question' flag and 'gender'
interruptions_question_analysis <- hansard_corpus %>%
  filter(gender %in% c("Male", "Female")) %>%
  group_by(gender, is_question = as.factor(question)) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    total_speeches = sum(total_speeches),
    interruption_proportion = total_interruptions / total_speeches,
    .groups = 'drop'
  )

# Print the resulting dataset to verify
print(interruptions_question_analysis)

```

```{r}
#| include: false
#| warning: false
#| message: false
# Filter and aggregate data based on the 'question' flag and 'gender'
interruptions_question_analysis <- hansard_corpus %>%
  filter(gender %in% c("Male", "Female")) %>%
  group_by(gender, is_answer = as.factor(answer)) %>%
  summarise(
    total_interruptions = sum(interject, na.rm = TRUE),
    total_speeches = n(),
    total_speeches = sum(total_speeches),
    interruption_proportion = total_interruptions / total_speeches,
    .groups = 'drop'
  )

# Print the resulting dataset to verify
print(interruptions_question_analysis)
```

```{r}
#| include: false
#| warning: false
#| message: false
# Assuming your dataset is named 'hansard_corpus'
# Calculate the word count for each 'body' entry
hansard_corpus1 <- hansard_corpus %>%
  mutate(word_count = str_count(body, "\\b\\w+\\b"))

# Print the data to see the word counts
head(hansard_corpus1)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-gov3
#| fig-cap: "Histogram of Word Counts in Speeches (Up to 250 Words)"

hansard_corpus1 <- hansard_corpus1 %>%
  filter(gender %in% fedchamb_flag == 0)

# Enhanced histogram with interjection data
ggplot(hansard_corpus1 %>% filter(word_count <= 250), aes(x = word_count, fill = as.factor(interject))) +
  geom_histogram(binwidth = 25, position = "identity", alpha = 0.6) +
  labs(
    title = "Histogram of Word Counts in Speeches (Up to 250 Words)",
    x = "Word Count",
    y = "Frequency",
    fill = "Interjection Occurred"
  ) +
  scale_fill_manual(values = c("0" = "steelblue", "1" = "salmon"), labels = c("No", "Yes")) +
  theme_minimal() +
  theme(legend.position = "top") +
  theme(
   plot.title = element_text(size = 14, face = "bold", hjust = 0.5))
```
@fig-gov3 shows word counts in speeches divided into two categories: speeches where an interjection occurred and those without any interruptions. We can see that speeches without interruptions (blue bars) are more frequent across all word count ranges, especially as the number of words increases. Conversely, interrupted speeches (red bars) are less frequent and generally shorter. This trend suggests that interruptions tend to happen in shorter speeches or possibly cause speakers to shorten their remarks. This graph gives a clear visual indication of how interruptions might disrupt or influence the length of parliamentary speeches.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-gov4
#| tbl-cap: "Gender-Based Interruptions and Word Count"
# Assuming hansard_corpus1 is already loaded and prepared
# First, filter to include only speeches up to 1000 words and only Male or Female
filtered_data <- hansard_corpus1 %>%
  filter(gender %in% c("Male", "Female"))

# Group by gender and whether there was an interjection
summary_table <- filtered_data %>%
  group_by(gender, interject) %>%
  summarise(
    num_speeches = n(),
    average_words = mean(word_count, na.rm = TRUE),
    min_words = min(word_count, na.rm = TRUE),
    max_words = max(word_count, na.rm = TRUE),
    .groups = 'drop'
  )

summary_table %>%
  gt()%>%
  tab_style(
    style = cell_text(align = "left"),  # Align text to the left
    locations = cells_body(columns = everything())  # Apply to all columns in the body
  ) %>%
  cols_label(
    gender = "Gender",
    interject = "Interruption",
    num_speeches = "Total Speeches",
    average_words = "Average Words",
    min_words = "Min Words",
    max_words = "Max Words")

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-gov5
#| fig-cap: "Gender-Based Interruptions and Word Count"
# Bar plot for total speeches with real numbers on the y-axis
ggplot(summary_table, aes(x = gender, y = num_speeches, fill = factor(interject))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("steelblue", "salmon"), labels = c("No", "Yes")) +
  scale_y_continuous(labels = scales::comma) +  # Display real numbers on the y-axis
  labs(
    title = "Total Speeches by Gender and Interruption",
    x = "Gender",
    y = "Total Speeches",
    fill = "Interruption"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

```
@tbl-gov4 and @fig-gov5 highlights that, on average, females tend to speak longer than males before an interruption occurs. For uninterrupted speeches, females average 560.72 words compared to 448.67 words for males. Even when interrupted, female speeches still have a higher average word count (37.76 words) than interrupted male speeches (28.76 words). This suggests that while females may face greater challenges overall, they tend to hold the floor longer before interruptions occur. However, the substantial drop in average word count for both genders during interruptions shows the disruptive impact, with females experiencing a steeper decline relative to their uninterrupted averages. This pattern indicates that while females often manage to speak at greater length initially, interruptions disproportionately diminish their overall speaking opportunities.

# Discussion {#sec-discussion}

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
git config --global http.postBuffer 524288000

git config --global http.postBuffer 629145600

git reset --soft HEAD~1

git lfs install
git lfs track "*.parquet"
git add .gitattributes
git commit -m "Track large parquet files with Git LFS"
git add data/01-raw_data/hansard-corpus/hansard_corpus_1998_to_2022.parquet
git commit -m "Add large Hansard Parquet file"
git push origin main

## Overview

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

\newpage

# References